---
title: "Multiple Linear Regression"
author: Ammar Mokhtar Gomaha Gaber
date: 20 January 2019
output: html_notebook
---

# Introductoion

# Setting the working directory
getwd()
# Read pollutants and meteorlogical data - daily (2003-2011)for model developing
```{r}
cm <- read.csv("Data/cm_2003-2015.csv", header=TRUE,sep=",")
names(cm)
```

Data summary to explore the data variables

```{r}
head(cm,5)
tail(cm,5)
```
# Convert wind speed unit by multiply by 0.277778
  
  
```{r}
cm$ws <- cm$WS*0.277778
names(cm)
```
```{r}
require(caTools)
sample <- sample.int(cm,SplitRatio=0.75)
train <- subset(cm, sample==TRUE)
test <- subset(cm, sample==FALSE)
```

# Arrange the data for regression analysis                # remove un desired columns
# Subset the3 data
# subset 2003-2011- dailly data for model development

```{r}
require(caTools)
sample <- sample.split(cm, SplitRatio=0.75)
train <- subset(cm,sample==TRUE)
test <- subset(cm, sample==FALSE)
```

Sampling
```{r}
samsize <- floor(0.75 *nrow(cm))
train_1 <- sample(seq_len(nrow(cm)), size=samsize)
train <- cm[train_1,]
test <- cm[-train_1,]

dim(train)
dim(test)
```

```{r}
# Impte the NA with mean for all variables
require(Hmisc)
train <- impute(train, fun=mean)
test <- impute(test, fun=mean)

train <- sapply(train, impute)
test <- sapply(test, impute)

# write the data to the current working directory
write.csv(train,"cm_train.csv")
write.csv(train,"cm_test.csv")
```







# Check the missing values

```{r}
table(is.na(cm))
sum(is.na(cm))
```

# Data Expration with ggplot
```{r}
install.packages("ggplot2")
  install.packages("grid")
  install.packages("gridExtra")
  
  install.packages(c("ggplot2","grid","gridExtra"))
  library(ggplot2)
  library(grid)
  library(gridExtra)
```
  

# Data exploration with ggplot

Plot Wind Speed in panel plot
remove missing values with na.omit=T
 
```{r}
train <- na.omit(train)
  par(mfrow=c(2,2))
  p1 <- ggplot(train, aes(x=ws, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Wind Speed")+ ylab("PM10 Concentration")+
    ggtitle("Wind Speed vs PM10 Concentrration")
```
 
 Plot Tmax
  
```{r}
 p2 <- ggplot(train, aes(x=tmax, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Maximum Temperature")+ ylab("PM10 Concentration")+
    ggtitle("Maximum Temperature vs PM10 Concentrration")
```
 
#Rainfall
```{r}
 p3 <- ggplot(train, aes(x=rain, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Rainfall")+ ylab("PM10 Concentration")+
    ggtitle("Rainfall vs PM10 Concentrration")
```
 
# Relative Hummidity
```{r}
 p4 <- ggplot(train, aes(x=rh, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Relative Humidity")+ ylab("PM10 Concentration")+
    ggtitle("Relative Humidity vs PM10 Concentrration")
```
 
  # Panel Plot
```{r}
par(mfrow=c(2,2))
  p1;p2;p3;p4
```
  
  
  # Pressure
  
```{r}
 ggplot(cm0311, aes(x=press, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Surface Pressure")+ ylab("PM10 Concentration")+
    ggtitle("Surface Pressure vs PM10 Concentrration")
```
 
  # Ozone
```{r}
 ggplot(cm0311, aes(x=o3, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Tropspheric Ozone")+ ylab("PM10 Concentration")+
    ggtitle("Troposheric Ozone vs PM10 Concentrration")
```
 
  #Carbon Monoxide
```{r}
ggplot(cm0311, aes(x=co, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Carbon Monoxide")+ ylab("PM10 Concentration")+
    ggtitle("CArbon Monoxide vs PM10 Concentrration")
```
  
  # Nox
```{r}
ggplot(cm0311, aes(x=nox, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Nox")+ ylab("PM10 Concentration")+
    ggtitle("Nitrogen Oxides vs PM10 Concentrration")
```
  
  # Sox
```{r}
ggplot(cm0311, aes(x=so2, y=pm10)) + geom_point(size=0.5, color="navy") + xlab("Sox")+ ylab("PM10 Concentration")+
    ggtitle("Sulphere Oxides vs PM10 Concentrration")
```
  

# Impute the missing values                                      

```{r}
table(is.na(cm$pm10)) 


cm$pm10.mean <- ifelse(is.na(cm1$pm10), mean(cm$pm10, na.rm = TRUE), cm$pm10)


table(is.na(cm$press)) 

cm$press.mean <- ifelse(is.na(cm), mean(cm, na.rm = TRUE), cm$press)
  
```
 
# check NA for all the data
```{r}
table(is.na(cm))
  
  cm$pm10.median <- ifelse(is.na(cm0311$pm10), median(cm0311$pm10, na.rm = TRUE), cm0311$pm10)
```
  
 
  #         Statistical Analysis  
  
  # Function to create histogram overlaped by density plot:
```{r}
library(ggplot2)
  
  tr <- function(a){
    ggplot(data = cm, aes(x=a,y=..density..))+ 
      geom_histogram(fill="blue", color="red", alpha=0.5, bins=100)+ geom_density()
    #ggplot(y)
  }
  tr(cm$tmax)
  
  cm$pm.trs <- (log(cm$pm10)+1)
  cm$pm10.trns <- log(cm$pm10)
  ggplot(cm, aes(x=pm10.trns,y=..density..))+ geom_histogram(fill="blue",color="black",alpha=0.5,bins=100)+geom_density()
  
  train <- cm0311
  test <- cm[1084:1563,-c(1,2,5,6,9:11)]
```
  
 
  #-----------------------------------------------------------------------------------------------------#
  # Calculate missing values percentage per column:
```{r}
mvtr <- sapply(train, function(x){sum(is.na(x)/length(x))})*100 
  mvts <- sapply(test, function(x){sum(is.na(x)/length(x))})*100  
  
  mvtr; mvts
```
  
  
  #----------------------------------------------------------------------------------------------------#
  #                                   MLR Model Constuction                                            
```{r}
 cm <- na.omit(cm) # using CM data from 2003 to 2011 for model parameters' estimation
```
 
  # Fit MLR model and generate the summary
  # correlation and plot of each explanatory variable with PM10
```{r}
 model.pm <-lm(pm10~ ., data = train[,-c(1,2)])
  summary(model.pm)
  
  pm10model        <- summary(lm(pm10~tmax+rh+rain+ws+press+o3+co+nox+so2, data=train))
  pmmodel.mean     <- summary(lm(pm10.mean~tmax+rh+rain+ws+press+o3+co+nox+so2, data=cm0311))
  pmmodel.median   <- summary(lm(pm10.median~tmax+rh+rain+ws+press+o3+co+nox+so2, data=cm0311))
  pm10model.trans  <- summary(lm(pm10.trns~tmax+rh+rain+ws+press+o3+co+nox+so2, data=train))
```
 
  # Multiple R-squared:  0.669,	Adjusted R-squared:  0.6659 
```{r}
 pm10model
```
  
     # Multiple R-squared:  0.6615,	Adjusted R-squared:  0.6584 
```{r}
pmmodel.mean  
```
  
  
     # Multiple R-squared:  0.6631,	Adjusted R-squared:  0.66
 
```{r}
pmmodel.median
```
 
    # Multiple R-squared:  0.628,	Adjusted R-squared:  0.6245 
```{r}
pm10model.trans
```
  y = a + b1*xi + b2*xi + ... + bn*xn
#-----------------------------------------#
```{r}
pm10 <- pm10model.trans$coefficients[1]+  pm10model.trans$coefficients[2] * tmax +  pm10model.trans$coefficients[3]*rh +
    pm10model.trans$coefficients[4]*rain +  pm10model.trans$coefficients[5]*ws +  pm10model.trans$coefficients[6]*press +
    pm10model.trans$coefficients[7]*o3 +  pm10model.trans$coefficients[8]*co+ pm10model.trans$coefficients[9]*so2
```
  
  
  # Model predoction
  
```{r}
pred.pm <- model.pm$coefficients[1]
  
  pred.pm <- 368.30188 + 0.52717 * tmax -0.63554*rh + 0.02717 * rain -
    0.69190 * ws + 0.68390 *o3 + 1.26036*nox + 41.22550 * com+
    4.27249 * so2
  
  pred.1 <- vector("numeric", length = 20)
  
    for (i in 1:20) {
      pred.1 [i]<- (368.30188 + 0.52717 *test$tmax[i] -0.63554*test$rh[i] + 0.02717 * test$rain[i] -
        0.69190 * test$WS[i] -0.34111* test$press + 0.68390 *test$o3[i] + 1.26036*test$nox[i] + 41.22550 * cm0311$co[i]+
        4.27249 * test$so2[i])
    }
```
  
 # Plot the model output
```{r}
plot(pred.1)
  summary(pred.1)
  str(pred.1)
  length(pred.1)
```
 
  
 
  #  Verification model results                                      
  
  #RMSE & MAE
  #Read stations data
  
```{r}
calb <- read.csv("F:/R_2018/An_Introduction_to_R_Course/data/Verification/Nan-2010.csv", header=T)
  names(calb)
  calb$WS <- calb$ws*0.277778
  calb[is.na(calb)]<- 0
  pmcal <- vector("numeric", 120)
  for (i in 1:120){
    pmcal[i] <- coef(pmmodel)[1,1]+coef(pmmodel)[2,1]*calb$tmax[i]+coef(pmmodel)[3,1]*calb$rh[i]+
      coef(pmmodel)[4,1]*calb$rain[i]+coef(pmmodel)[5,1]*calb$ws[i]+coef(pmmodel)[6,1]*calb$press[i]+
      coef(pmmodel)[7,1]*calb$o3[i]+coef(pmmodel)[8,1]*calb$co[i]+coef(pmmodel)[9,1]*calb$nox[i]+
      coef(pmmodel)[10,1]*calb$so2[i]
  }
  
  for (i in 1:120) {
    pmcal[i]<- (355.5125 + 355.5125 * calb$tmax -0.6644001 * calb$rh + 0.04429299 * calb$rain -0.7839519 * calb$ws +-
      0.3243547 * calb$Press + 0.6768686 * calb$o3+ 39.40836 * calb$co + 1.283827 * calb$nox + 4.648964 * calb$so2)
    
  }
```
  
  
  
  #head(pmcal)
  #write.table(pmcal,paste(calb.files[j],".txt")) # prediction values
  write.table(pmcal,paste(calb,".txt")) # prediction values
  write.table(pmcal,paste("L2015c.txt"))
  # Replace negative values with zeros
  pmcal[pmcal<0] <- 0* (pmcal[pmcal<0])
  #------------------------------------------------------------------#
  # Install packages
```{r}
install.packages("Metrics")
  library(Metrics)
```
  
  
  #Compute Root Mean Square Error
```{r}
rmse(calb$pm10[1:120], pmcal[1:120])
```
  
  
  #Compute  Mean Absolute Error
```{r}
mae(calb$pm10[1:120], pmcal[1:120])
```
  
  # IOA 
```{r}
install.packages("hydroGOF")
  library(hydroGOF)
  d(calb$pm10, pmcal)
  plot(calb$pm10)
```
  
  # F-test 
```{r}
 f_test <- var.test(calb$pm10,pmcal, conf.level = 0.95)
  f_test
```
 
  # t-test 
```{r}
t_test <- t.test(calb$pm10,pmcal,conf.level = 0.95)
  t_test
```
  
  
  