---
title: "Regression Model: Construction,Validation and Verification"
author: "Ammar Mokhtar Gomah Gaber"
date: "January 2019"
output:
  html_document:
    df_print: paged
---

#Introduction to Predictive Models using Various Methods
The following examples we will use Big Mart Sales data set:
1. Graphical Representation of Variables
ggplot2 package will be used to create the graphs. The graphs will allsow us to uns=derstand the distribution and frequency of variables in the data set
```{r}
train <- read.csv("F:/R_2019/R_Course_Part_II/Modules/Data/Big_Mart_Sales/Train.csv")
test <- read.csv("F:/R_2019/R_Course_Part_II/Modules/Data/Big_Mart_Sales/test.csv", header = TRUE)
```


```{r}
library(ggolpt2)
# Plot
ggolpt(train, aes(x=Item_Visibility, y=Item_Outlet_Sales))+ geom_point(size=2.5,
                                                                       color="navy")+ xlab("Item Visibility")+ ylab("Item Outlet Sales")+ ggtitle("Item Visibility vs Item Outlet Sales")
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
