---
title: "An Introduction to Ensemble Predictives Models"
author: "Ammar Mokhtar Gomaha Gaber"
date: "January 2019"
output:
  html_document:
    df_print: paged
---

# Ensemble Predictive Models
A model is usually a predictive model or a model of the structure if the data that we want ot construct or discover from the data set, such as a decsion tree, a neural network, a support vector maachine, etc.
The process of generating models from data is called $learning$ or training, which is accomplished by learning algorithm. There are differnt learning settings, among which the most common ones are $supervised learning$ and $unsupervised learning$. 
# Example
Install devtools library and medley from GitHub
```{r}
library(devtools)
install_github(c("medley","mewo2"))
library(medley)
```



```{r}

# Generate some data
y <- rnorm(1000)       # predectand
x1 <- (y > 0) ; x2 <- y * rnorm(1000)
x3 <- rnorm(1000, mean=y, sd=1) ; x4 <- (y > 0)& (y < 3)
x5 <- rbind(1000, size=4, prob=exp(y)/(1+exp(y)))
x6 <- (y < -2)| (y > 2)
# make data frame
data <- data.frame(y=y,x1=x1,x2=x2,x3=x3,x4=x4,x5=x5,x6=x6)
# prepare training and test data sets
train <- sample(1:1000, size = 500)
trainData <- data[train,]
testData <- data[-train]
```
# Basic Model
Generate linear modelwith lm()function and test ti with the new data det (testData) using the lowest value of the RMSE. The samllest the value the better.

```{r}
library(tree)
lm1 <- lm(y ~ ., data=trainData)
rmse(predict(lm1, data=testData), testData$y)
```

# Tree MOdel
Generate another model using tree package and check its accuracy using the same statistical perfromance indicator (RMSE).

```{r}
tree1 <- tree(y ~ ., data=trainData)
rmse(predict(tree1,data=testData), testData$y)
```
# Third model
Generate a third model using tree.
```{r}
tree2 <- tree(y ~ ., data=trainData[sample(1:dim(trainData)[1]),])
```

# Comibning the Models
Combine the the lm1 and tree1 models
```{r}
combine1 <- predict(lm1, data=testData)/2 + predict(tree1, data=testData)/2
rmse(combine1,testData$y)
```
Combine l,1 and tree1 and tree2 models
```{r}
combine2 <- (predict(lm1, data=testData)/3 + predict(tree1, data=testData)/3 +
  predict(tree2, data=testData)/3)
rmse(combine2, testData$y)
```
The error of prediction is getting smaller.

# Medley PAckage
This package is capable to perfrom an ensamble models
```{r}
library(devtools)
instal.github("medley","mewo2")
library(medley)
library(e1071)                     # use SVM
library(rndomForest)
# crreate predictors from the training data set.
x <- trainData[,-1]
# crreate predctand from the training data set.
y <- trainData$y
# crreate predictors from the training data set.
newx <- testData[,-1]
```

This technique is most used in Kaggle competitions (http://www.kaggle.com/users/10748/martin-o-leary)

# Blending Models (part 1)
Now we can blend all the models using the medley apckage:
```{r}
m <- create.medley(x,y,errfunc=rmse);
for (g in 1:10) {
  m <- add.medley(m, svm, list(sigma=1e-3 *g));
}
```

This function creats 10 models using the SUpport Vector Machine (SVM)with 10 different values of gamma. The caculating time and error is also calcualted for each model.
# Blending Models (part 2)
medley() function can also be used to generate ensemble models using randomForest():
```{r}
for (mt in 1:2) {
  m <- add.medley(m, randomForest, list(mtry=mt));
}
```

Two models will be generated with differnt mt.Calculated time and error values will be produced also.
The function can identify the good mode

```{r}
m <- prune.medley(m, 0.8);
rmse(predict(m,newx), testData$y)
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
